---
title: "Introduction to gaptrace"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to gaptrace}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

`gaptrace` transforms tabular sustainability data collection trackers into executive-level visual narratives. This vignette demonstrates the complete workflow from data ingestion to exporting operational target lists.

## 1. Load Dummy Data

The package includes a synthetic dataset named `sustain_dummy_data` representing ~50 rows of simulated environmental consulting data.

```{r setup}
library(gaptrace)

# View the first few rows
head(sustain_dummy_data)
```

## 2. Ingest Data

First, we ingest the raw data using `sustain_ingest()`. This function validates the schema and formats columns.

```{r ingest}
tracker <- sustain_ingest(sustain_dummy_data)
class(tracker)
```

## 3. Visualize Progress

We can plot an executive waffle chart showing the proportion of data collected versus missing.

```{r waffle, fig.width=8, fig.height=5}
plot_waffle_progress(tracker)
```

## 4. Visualize Interdependencies

We can also generate an interdependency network graph linking specific data points to overarching themes. This can highlight how missing data nodes act as bottlenecks for central themes.

```{r network, fig.width=8, fig.height=6}
# Generate a static network graph
plot_theme_network(tracker, interactive = FALSE)
```

*(Note: Use `interactive = TRUE` to render an interactive `visNetwork` graph in your local R environment.)*

## 5. Export Targeted Action List

Finally, we export a prioritized, filtered `.xlsx` list to share with operational teams.

```{r export, eval=FALSE}
# Create an Excel file containing only 'Not Delivered' items
export_targeted_list(
  tracker, 
  filter_status = c("Not Delivered"), 
  file_path = "urgent_action_list.xlsx"
)
```
